Polymethylation scores results

### 2021-10-28 Reports/Results.Rmd
rewriting results to include/exclude ewas results

library(ggplot2)
library(kableExtra)
library(ggpubr)
library(tidyr)
library(stringr)
library(dplyr)
library(compareGroups)
library(kableExtra)
library(gtsummary)
library(gtools)
library(haven)
library(sjlabelled)
library(matchmaker)
library(EpiDISH)
library("PerformanceAnalytics")
library(DescTools)
library(purrr)
library(broom)
library(broom.mixed)
library(ggplot2)
library(multcomp)
library(ggpubr)
library(cowplot)
library(geepack)
library(lme4)
library(pROC)
library(lmerTest)
library(stringr)
library(sva)
library(gtsummary)
library(gt)
library(here)
library(DiagrammeRsvg)
library(xml2)
library(ggrepel)
```

@@

all_methyl=c("globalmethylation", "pediatric", "grim", "Epithelial.cells_saliva", "Leukocytes_saliva", "SSnewborn_center", "SSnewbornCT_center", "SSolder_center", "cg05575921")
main_methyl=c("globalmethylation", "pediatric", "SSnewbornCT_center", "cg05575921")

@@

all_methyl_levels=c("SSnewbornCT_center", "SSnewborn_center",  "SSolder_center",
                                       "pediatric" , "grim", "cg05575921", 
                                      "globalmethylation","Epithelial.cells_saliva", "Leukocytes_saliva")

@@

maternaldata=myFF %>% filter(!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) & PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth!='Missing')  
n_withmaternaldata=paste0('m=', maternaldata %>% nrow(), ' samples & n=', maternaldata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with maternal covariate data')

@@

n_Child<-paste0('n=', childdata %>%filter(childteen=='C')  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata%>%filter(childteen=='C' & k5f1l=='2 no')%>% nrow(), 
                           " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', childdata%>%filter(k5f1l!= "2 no" & childteen=='C')%>%nrow(), 
                             " individuals smoking \n or missing child smoking data \n at 9 year visit")
n_Teen<-paste0('n=', childdata %>%filter(childteen=='T')   %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata %>%filter(childteen=='T' & k6d40=="2 No") %>% nrow(), 
                          " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', childdata %>% filter(k6d40!="2 No" & childteen=='T') %>% nrow,
                            " individuals smoking \n or missing focal child smoking data \n at 15 year visit") 
if(childdata %>% filter((childteen=='C' & k5f1l=='2 no') | (childteen=='T' & k6d40=='2 No'))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

@@

graph1

@@

main_fig=main_fig+1            
```

@@

exclusion_tbl

@@

@@

```{r, fig.width=12, fig.height=12}
library(Hmisc)
library(corrplot)
res2<-modeldata%>%dplyr::select(any_of(all_methyl))%>%cor()
#labels
colnames(res2)=all_methyl_labels
rownames(res2)=all_methyl_labels
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)

@@

@@

visit_cor=age_wide_methyl%>%
  ggplot(aes(x=`Age 9`, y=`Age 15`))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+
  scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))

visit_cor

@@

@@

```{r, fig.height=11}
dist_methyl=age_long_methyl%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, color=childteen))+
  geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..", label.y.npc=0.87)+
  scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  geom_point(alpha=0)+geom_line(alpha=0, aes(group=idnum))+geom_smooth(aes(group=1), method='lm', color='black')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))+
  theme(legend.position = 'bottom')

dist_methyl

@@

@@

fig2

@@

@@

```{r bivariate table}
smoking_tbl=modeldata %>% 
  dplyr::select(smkPreg_binary, m1g2_YesNoPreg, m1g3_YesNoPreg, cm1inpov, PostnatalMaternalSmokingAny,
         SmkAtVisitPastmonth, ancestry, cm1bsex, globalmethylation, pediatric, Epithelial.cells_saliva,
         Leukocytes_saliva, SSnewbornCT_center, SSolder_center, cg05575921, childteen)%>%
  tbl_strata(strata=childteen, .tbl_fun=~.x %>%
               tbl_summary(by=smkPreg_binary, statistic=
                             list(all_continuous()~"{mean} ({sd})"))%>%
               add_n()%>%add_p(test=list(all_continuous()~"t.test")))

@@

smoking_tbl

@@

@@

```{r, fig.width=12}
my_outcomes=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)')

@@

forest_plot

@@

@@

```{r, fig.width=14}
with(modeldata, table(ancestry, PostnatalMaternalSmokingAny, smkPreg_binary))%>%kbl()

@@

forest_plot_compareModels

@@

@@

```{r, fig.width=14, fig.height=10}

@@

modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', 'PostnatalMaternalSmokingAny', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))%>%

  mutate(exposed=paste0(smkPreg_binary, ' prenatal smoke exposed & ', PostnatalMaternalSmokingAny))%>% 
  filter(Methyl_factor%in% main_methyl[3:4])%>%
  ggplot(aes(x=exposed, y=`Methylation summary measure`, color=exposed))+
    geom_violin()+geom_boxplot(width=0.2, position = position_dodge(0.9))+
  stat_summary(fun.data = give.n, geom = "text", fun.y = max,
                  position = position_dodge(width = 0.75))+
  stat_compare_means(comparisons=my_comparisons)+
    facet_grid(Methyl_factor~childteen, scales='free',labeller = labeller(Methyl_factor=all_methyl_labels))+
    scale_x_discrete(labels=function(x) str_wrap(x, width = 10))+
    theme(legend.position = 'right')+
  guides(color=guide_legend(nrow=4,byrow=TRUE))
sup_fig=sup_fig+1
```

@@

regression_table_age

@@

@@

regression_table_ancestry

@@

@@

```{r, fig.width=12, fig.height=8}
#bring the heatmap function
covar=c('Sample_Plate', 'Slide', 'Array', 'cm1bsex', 'smkPreg_binary', 'childteen', 'ancestry')

@@

bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))

@@

@@

regression_table_ancestry_svs

@@

@@

```{r, fig.width=10}

@@

auc_fig

@@

@@

roc_table1%>%mutate(`pval comparison to sustained smoking score (newborns)`=map_dbl(roctest, ~round(.x$p.value, 4)))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to sustained smoking score (newborns)`)%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs polymethylation score, newborns, cell type controlled)'))
```

@@

roc_table2%>%mutate(`pval comparison to base model`=map_dbl(roctest, ~round(.x$p.value, 4)))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to base model`)%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs base w/ no methyl)'))

@@

@@

@@

```{r}
child_m1=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'child1_TopHits.RDS'))
teen_m1=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'teen1_TopHits.RDS'))
```

```{r}
#lme_base=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'lme_ewas_base.RDS'))
#lme_smk = lme_base %>% filter(term=='smkPreg_binaryYes')%>%mutate(p.bh=p.adjust(p.value, 'hochberg'))
#lme_smk=add_annotations(lme_smk, type='lme')
lme_smk=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'lme_smoking.RDS'))
#mutate columns to get matching columns in lme_smk as in child & teen _m1
lme_smk = lme_smk %>% mutate(logFC=estimate, CI.L=conf.low, CI.R=conf.high, P.Value=p.value, adj.P.Val=p.bh)
```

@@

par(mfrow=c(3, 1), mar=c(5.1, 4.1, 5.1, 2.1))
  qqman::qq(child_m1$P.Value, main='Age 9: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(child_m1), 3)), side=3)

@@

sup_fig=sup_fig+1

@@

```{r, cache=T}
library(tibble)
base_models=list('Age_9'=child_m1%>%rownames_to_column(var='CPG'), 'Age_15'=teen_m1%>%rownames_to_column(var='CPG'), "Mixed_effect_model"=lme_smk)%>%Map(cbind, ., ModelType=names(.))

@@

@@

```{r}
library(ggupset)
sig_cpg_list<-map(base_models, ~.x %>%filter(P.Value<10e-4)%>%dplyr::select(CPG, ModelType))%>%
  bind_rows()%>%
  group_by(CPG)%>%summarise(Models=list(ModelType))

@@

```{r, fig.width=14}
volcs=ggarrange(
ggplot(child_m1%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
ggplot(teen_m1%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
ggplot(lme_smk%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
labels=c('A', 'B', 'C'), 
common.legend = TRUE, 
nrow=1
)

ggexport(volcs, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')))
ggexport(volcs, filename=file.path(outdir, paste0('Figure', main_fig, '.png')))

volcs
```

@@

```{r}
all_gaps=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'gaphunter_gaps.RDS'))
```

@@

```{r, fig.width=14}
topCPGs<-read.csv(paste0(datadir, "/TopCpGsWiklundEtAl2019.csv"))
topCPGs<-topCPGs %>% mutate(estimate=as.numeric(gsub("−\\s", "-", gsub(" .*", "", β..SE.))), std.error=as.numeric(sub("\\).*", "", sub(".*\\(", "", β..SE.))), p.value=P.value, source="Wiklund et al. 2019", conf.low=(estimate-1.96*std.error), conf.high=(estimate+1.96*std.error), set ='')%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)

@@

```{r, fig.width=14}
comparison=left_join(
  lme_smk%>%dplyr::select(CPG, gene, estimate, p.value)%>%dplyr::rename_with(~paste0('FF_', .x), .cols=c(estimate, p.value)), 
  cpgs%>%filter(set=='SSnewbornCT')%>%mutate(CPG=CpG)%>%dplyr::select(CPG, Gene, estimate, p.value)%>%dplyr::rename_with( ~paste0('Joubert_', .x), .cols=c(estimate, p.value))
  )%>%filter(!is.na(Joubert_estimate))

@@

comparison %>% mutate(s1=sign(FF_estimate), s2=sign(Joubert_estimate), agreement=case_when(s1==s2~'Direction consistent', s1<s2~'FF -, Joubert +', s1>s2~'FF+, Joubert-'))%>%group_by(agreement)%>%summarise(n=n())

@@

comp_joubert=ggarrange(comp_nofilter, comp_filter, common.legend = F, labels=c('A', 'B'))
ggexport(comp_joubert, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.png')))

comp_joubert
```

@@

```{r}
top_cpgs = map(base_models, ~.x %>%filter(adj.P.Val<0.05)%>%dplyr::select(CPG, ModelType, gene))%>%bind_rows()%>%mutate(label=paste0(gene, ': ', CPG))
top_cpg_gaps=top_cpgs[top_cpgs$CPG %in% rownames(all_gaps$proberesults),]
top_cpgs=top_cpgs[!(top_cpgs$CPG %in% rownames(all_gaps$proberesults)),]
top_cpg_groups=top_cpgs %>%group_by(CPG)%>%dplyr::summarise(models=list(ModelType))
top_cpg_names=c(top_cpgs%>%pull(label)%>%unique(), 'AHRR: cg05575921', 'MYO1G: cg22132788')
top_cpgs=c(top_cpgs %>% pull(CPG)%>%unique(), 'cg05575921', 'cg22132788')

@@

```{r, fig.width=14}

@@

@@

ggexport(plot_sharedCpG, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))

plot_sharedCpG

>>>>

library(ggplot2)
library(kableExtra)
library(ggpubr)
library(tidyr)
library(stringr)
library(dplyr)
library(compareGroups)
library(kableExtra)
library(gtsummary)
library(gtools)
library(haven)
library(sjlabelled)
library(matchmaker)
library(EpiDISH)
library("PerformanceAnalytics")
library(DescTools)
library(purrr)
library(broom)
library(broom.mixed)
library(ggplot2)
library(multcomp)
library(ggpubr)
library(cowplot)
library(geepack)
library(lme4)
library(pROC)
library(lmerTest)
library(stringr)
library(sva)
library(gtsummary)
library(gt)
library(here)
library(DiagrammeRsvg)
library(xml2)
library(ggrepel)

include_ewas=F
stand_alone=T
```

@@

all_methyl=c("globalmethylation", "pediatric", "grim", 'DNAmPACKYRS', "Epithelial.cells_saliva", "Leukocytes_saliva", "SSnewborn_center", "SSnewbornCT_center", "SSolder_center", "cg05575921")
main_methyl=c("globalmethylation", "pediatric", "SSnewbornCT_center", "cg05575921")

@@

all_methyl_levels=c("SSnewbornCT_center", "SSnewborn_center",  "SSolder_center",
                                       "pediatric" , "grim", 'DNAmPACKYRS',  "cg05575921", 
                                      "globalmethylation","Epithelial.cells_saliva", "Leukocytes_saliva")

@@

maternaldata=myFF %>% filter(!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) & PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth!='Missing')

n_withmaternaldata=paste0('m=', maternaldata %>% nrow(), ' samples & n=', maternaldata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with maternal covariate data')

@@

n_Child<-paste0('n=', childdata %>%filter(childteen=='C')  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata%>%filter(childteen=='C' & k5f1l=='2 no')%>% nrow(), 
                           " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', childdata %>% filter((k5f1l=="1 yes" | is.na(k5f1l)) & childteen=='C') %>% nrow(), 
                             " individuals smoking \n or missing child smoking data \n at 9 year visit")
n_Teen<-paste0('n=', childdata %>%filter(childteen=='T')   %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata %>%filter(childteen=='T' & k6d40=="2 No") %>% nrow(), 
                          " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', childdata %>% filter((k6d40=="1 Yes" | is.na(k6d40) | k5f1l=='1 yes') & childteen=='T') %>% nrow,
                            " individuals smoking \n or missing focal child smoking data \n at 15 year visit")

if(childdata %>% filter((childteen=='C' & k5f1l=='2 no') | (childteen=='T' & k6d40=='2 No' & k5f1l!='1 yes'))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

@@

@@

main_fig=main_fig+1            
```
```{r, eval=stand_alone}
graph1
```

@@

@@

```{r, eval=stand_alone}
exclusion_tbl
```

```{r wide data}
modeldata_wide = modeldata%>%
  mutate(childteen=gsub(' ', '_', childteen))%>%
  dplyr::select(any_of(c('idnum', 'ancestry', 'childteen', 'cm1bsex', 'ancestry', 
                         'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'cm1inpov',
                         'smkPreg_binary',  all_methyl)))%>%
  pivot_wider(names_from = childteen, values_from=any_of(all_methyl))

model_twovisit=modeldata_wide[complete.cases(modeldata_wide), ]
```

@@

```{r}
library(Hmisc)
library(corrplot)
res2<-modeldata%>%dplyr::select(any_of(all_methyl))%>%cor()
#labels
colnames(res2)=all_methyl_labels
rownames(res2)=all_methyl_labels

@@

```{r, eval=stand_alone, fig.width=12, fig.height=12}
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)

@@

visit_cor=age_wide_methyl%>%
  ggplot(aes(x=`Age 9`, y=`Age 15`))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+
  scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), ncol=3)

@@

```{r, eval=stand_alone, fig.width=12, fig.height=12}
visit_cor
```

@@

```{r}
dist_methyl=age_long_methyl%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, color=childteen))+
  geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..", label.y.npc=0.87)+
  scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  geom_point(alpha=0)+geom_line(alpha=0, aes(group=idnum))+geom_smooth(aes(group=1), method='lm', color='black')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), ncol=3)+
  theme(legend.position = 'bottom')

@@

```{r, eval=stand_alone, fig.height=11}
dist_methyl
```

@@

@@

```{r, eval=stand_alone}
fig2
```

@@

```{r bivariate table}
smoking_tbl=modeldata %>% 
  dplyr::select(smkPreg_binary, m1g2_YesNoPreg, m1g3_YesNoPreg, cm1inpov, PostnatalMaternalSmokingAny,
         SmkAtVisitPastmonth, ancestry, cm1bsex, globalmethylation, pediatric, grim, DNAmPACKYRS, Epithelial.cells_saliva,
         Leukocytes_saliva, SSnewbornCT_center, SSolder_center, cg05575921, childteen)%>%
  tbl_strata(strata=childteen, .tbl_fun=~.x %>%
               tbl_summary(by=smkPreg_binary, statistic=
                             list(all_continuous()~"{mean} ({sd})"))%>%
               add_n()%>%add_p(test=list(all_continuous()~"t.test")))

@@

@@

```{r, eval=stand_alone}
smoking_tbl
```

@@

```{r}
my_outcomes=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)')

@@

@@

```{r, eval=stand_alone, fig.width=12}
forest_plot
```

@@

```{r}
with(modeldata, table(ancestry, PostnatalMaternalSmokingAny, smkPreg_binary))%>%kbl()

@@

@@

```{r, eval=stand_alone, fig.width=14}
forest_plot_compareModels
```

@@

@@

dna_prepost=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', 'PostnatalMaternalSmokingAny', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))%>%

  mutate(exposed=paste0(smkPreg_binary, ' prenatal smoke exposed & ', PostnatalMaternalSmokingAny))%>% 
  filter(Methyl_factor%in% main_methyl[3:4])%>%
  ggplot(aes(x=exposed, y=`Methylation summary measure`, color=exposed))+
    geom_violin()+geom_boxplot(width=0.2, position = position_dodge(0.9))+
  stat_summary(fun.data = give.n, geom = "text", fun.y = max,
                  position = position_dodge(width = 0.75))+
  stat_compare_means(comparisons=my_comparisons)+
    facet_grid(Methyl_factor~childteen, scales='free',labeller = labeller(Methyl_factor=all_methyl_labels))+
    scale_x_discrete(labels=function(x) str_wrap(x, width = 10))+
    theme(legend.position = 'right')+
  guides(color=guide_legend(nrow=4,byrow=TRUE))

ggexport(dna_prepost, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=10, height=10)
sup_fig=sup_fig+1
```

```{r, eval=stand_alone, fig.width=14, fig.height=10}
dna_prepost
```

@@

@@

```{r, eval=stand_alone}
regression_table_age
```

@@

@@

```{r, eval=stand_alone}
regression_table_ancestry
```

@@

```{r}
#bring the heatmap function
covar=c('Sample_Plate', 'Slide', 'Array', 'cm1bsex', 'smkPreg_binary', 'childteen', 'ancestry')

@@

@@

```{r, eval=stand_alone, fig.width=12, fig.height=8}
bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))
```

@@

@@

```{r, eval=stand_alone}
regression_table_ancestry_svs
```

@@

@@

@@

```{r, eval=stand_alone, fig.width=10}
auc_fig
```

@@

rct1=roc_table1%>%mutate(`pval comparison to sustained smoking score (newborns)`=map_dbl(roctest, ~round(.x$p.value, 4)))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to sustained smoking score (newborns)`)

```{r, eval=stand_alone}
rct1%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs polymethylation score, newborns, cell type controlled)'))
```

@@

rct2=roc_table2%>%mutate(`pval comparison to base model`=map_dbl(roctest, ~round(.x$p.value, 4)))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to base model`)

rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs base w/ no methyl)'))

@@

```{r, eval=stand_alone}
rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value (vs base w/ no methyl)'))
```

@@

```{r, eval=include_ewas}

@@

```{r, eval=include_ewas}
child_m1=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'child1_TopHits.RDS'))
teen_m1=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'teen1_TopHits.RDS'))
```

```{r, include=include_ewas}
#lme_base=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'lme_ewas_base.RDS'))
#lme_smk = lme_base %>% filter(term=='smkPreg_binaryYes')%>%mutate(p.bh=p.adjust(p.value, 'hochberg'))
#lme_smk=add_annotations(lme_smk, type='lme')
lme_smk=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'lme_smoking.RDS'))
#mutate columns to get matching columns in lme_smk as in child & teen _m1
lme_smk = lme_smk %>% mutate(logFC=estimate, CI.L=conf.low, CI.R=conf.high, P.Value=p.value, adj.P.Val=p.bh)
```

@@

```{r, eval=stand_alone, eval=include_ewas}
par(mfrow=c(3, 1), mar=c(5.1, 4.1, 5.1, 2.1))
  qqman::qq(child_m1$P.Value, main='Age 9: Base model EWAS')
  mtext(paste0(expression(lambda), '=', round(calcLambda(child_m1), 3)), side=3)

@@

@@

```{r, cache=T, eval=include_ewas}
library(tibble)
base_models=list('Age_9'=child_m1%>%rownames_to_column(var='CPG'), 'Age_15'=teen_m1%>%rownames_to_column(var='CPG'), "Mixed_effect_model"=lme_smk)%>%Map(cbind, ., ModelType=names(.))

@@

```{r, eval=stand_alone, eval=include_ewas}
beta_compare
```

@@

```{r, eval=include_ewas}
library(ggupset)
sig_cpg_list<-map(base_models, ~.x %>%filter(P.Value<10e-4)%>%dplyr::select(CPG, ModelType))%>%
  bind_rows()%>%
  group_by(CPG)%>%summarise(Models=list(ModelType))

@@

```{r, eval=include_ewas}
volcs=ggarrange(
ggplot(child_m1%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
ggplot(teen_m1%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
ggplot(lme_smk%>%mutate(adjPval=ifelse(adj.P.Val<0.05, '<0.05', ifelse(adj.P.Val<0.2, '<0.2', '>0.2'))), aes(x=logFC, y=-log10(P.Value), color=adjPval))+geom_point()+scale_x_continuous('Beta (regression coefficient)')+
  scale_color_manual('Benjamini-Hochberg adjusted P value', values=c('red', 'orange', 'black'))+
  scale_y_continuous(limits=c(0, 9), breaks=c(0, 2, 4, 6, 8))+
  scale_x_continuous(limits=c(-0.06, 0.06), breaks=c(-0.05, 0, 0.05)),
labels=c('A', 'B', 'C'), 
common.legend = TRUE, 
nrow=1
)

ggexport(volcs, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')))
ggexport(volcs, filename=file.path(outdir, paste0('Figure', main_fig, '.png')))
```

```{r, eval=stand_alone, fig.width=14, }
volcs
```

@@

```{r, eval=include_ewas}
all_gaps=readRDS(file.path(datadir, 'CreatedData', 'EWAS', 'gaphunter_gaps.RDS'))
```

@@

```{r, eval=include_ewas}
topCPGs<-read.csv(paste0(datadir, "/TopCpGsWiklundEtAl2019.csv"))
topCPGs<-topCPGs %>% mutate(estimate=as.numeric(gsub("−\\s", "-", gsub(" .*", "", β..SE.))), std.error=as.numeric(sub("\\).*", "", sub(".*\\(", "", β..SE.))), p.value=P.value, source="Wiklund et al. 2019", conf.low=(estimate-1.96*std.error), conf.high=(estimate+1.96*std.error), set ='')%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)

@@

```{r, fig.width=14, eval=include_ewas}
comparison=left_join(
  lme_smk%>%dplyr::select(CPG, gene, estimate, p.value)%>%dplyr::rename_with(~paste0('FF_', .x), .cols=c(estimate, p.value)), 
  cpgs%>%filter(set=='SSnewbornCT')%>%mutate(CPG=CpG)%>%dplyr::select(CPG, Gene, estimate, p.value)%>%dplyr::rename_with( ~paste0('Joubert_', .x), .cols=c(estimate, p.value))
  )%>%filter(!is.na(Joubert_estimate))

@@

#comparison %>% mutate(s1=sign(FF_estimate), s2=sign(Joubert_estimate), agreement=case_when(s1==s2~'Direction consistent', s1<s2~'FF -, Joubert +', s1>s2~'FF+, Joubert-'))%>%group_by(agreement)%>%summarise(n=n())

@@

comp_joubert=ggarrange(comp_nofilter, comp_filter, common.legend = F, labels=c('A', 'B'))
ggexport(comp_joubert, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.png')))
```

```{r, eval=stand_alone, fig.width=14}
comp_joubert
```

@@

```{r, eval=include_ewas}
top_cpgs = map(base_models, ~.x %>%filter(adj.P.Val<0.05)%>%dplyr::select(CPG, ModelType, gene))%>%bind_rows()%>%mutate(label=paste0(gene, ': ', CPG))
top_cpg_gaps=top_cpgs[top_cpgs$CPG %in% rownames(all_gaps$proberesults),]
top_cpgs=top_cpgs[!(top_cpgs$CPG %in% rownames(all_gaps$proberesults)),]
top_cpg_groups=top_cpgs %>%group_by(CPG)%>%dplyr::summarise(models=list(ModelType))
top_cpg_names=c(top_cpgs%>%pull(label)%>%unique(), 'AHRR: cg05575921', 'MYO1G: cg22132788')
top_cpgs=c(top_cpgs %>% pull(CPG)%>%unique(), 'cg05575921', 'cg22132788')

@@

```{r, fig.width=14, eval=include_ewas}

@@

```{r, eval=stand_alone, fig.width=14}
ancestry_differences
```

```{r, eval=include_ewas}

@@

ggexport(plot_sharedCpG, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
```

```{r, eval=stand_alone, fig.width=14}
plot_sharedCpG
```
